---
title: "Alliances Analysis"
output: html_notebook
---


```{r}
library(ggplot2)
library(grid)
library(gridExtra)
library(tidyr)
library(dplyr)
library(viridis)
library(oce)
library(vcd)
library(vcdExtra)
library(extracat)

## Set file path
path = "/Users/cynthiaclement/Desktop"



################################# Looking at US Alliances #################################
dyad_al= read.csv(paste(path, "version4.1_csv/alliance_v4.1_by_dyad.csv", sep= "/"))
dyad_al_year = read.csv(paste(path, "version4.1_csv/alliance_v4.1_by_dyad_yearly.csv", sep= "/"))

member_alliances = read.csv(paste(path, "version4.1_csv/alliance_v4.1_by_member.csv", sep= "/"))
member_al_year = read.csv(paste(path, "version4.1_csv/alliance_v4.1_by_member_yearly.csv", sep= "/"))

dir_al_year= read.csv(paste(path, "version4.1_csv/alliance_v4.1_by_directed_yearly.csv", sep= "/"))
dir_al = read.csv(paste(path, "version4.1_csv/alliance_v4.1_by_directed.csv", sep= "/"))


us_alliances <- dyad_al[dyad_al$ccode1 == 2,] 
alliances <- us_alliances[,c(2:5, 8, 11)]
alliances<- gather(alliances, st_end, year, 	dyad_st_year:dyad_end_year)
ggplot(alliances, aes(x=year, y = state_name2, color = st_end)) + geom_point() 




```



alliance v4 by member.* : this file contains one observation for every state in an alliance at the time of their alliance initiation  
alliance v4 by member yearly.* : this file expands the alliance initiation format to include one observation for each state-year of the alliance  
alliance v4 by dyad.* : the file contains one observation for every dyad in an alliance at the time of their alliance initiation  
alliance v4 by dyad yearly.* : this file expands the dyadic alliance data to include one observation for each allied dyad-year  
alliance v4 by directed.* : this file contains one observation for each directed dyad in an alliance at the time of their alliance initiation    
alliance v4 by directed yearly.* : this file expands the directed-dyad alliance data to include one observation for each allied directed dyad-year  

alliance_v4.1_by_directed_yearly.csv
alliance_v4.1_by_directed.csv
alliance_v4.1_by_dyad_yearly.csv
alliance_v4.1_by_dyad.csv
alliance_v4.1_by_member_yearly.csv
alliance_v4.1_by_member.csv




#WWI

```{r}
allied <- c("United States of America", "United Kingdom", "Russia", "Japan", "Italy")
allied_ccode <- member_alliances$ccode[match(allied, member_alliances$state_name)]

central <- c("Germany", "Turkey", "Austria-Hungary", "Romania", "Bulgaria")
central_ccode <- member_alliances$ccode[match(central, member_alliances$state_name)]
```


#WWII

```{r}
allies <- c("United States of America", "United Kingdom", "France", "Russia", "Australia", "Belgium", "Brazil", "Canada", "China", "Denmark", "Greece", "Netherlands", "New Zealand", "Norway", "Poland", "South Africa", "Yugoslavia")
allies_ccode <- member_alliances$ccode[match(allies, member_alliances$state_name)]

axis <- c("Germany", "Italy", "Japan", "Hungary", "Romania", "Bulgaria")
axis_ccode <- member_alliances$ccode[match(axis, member_alliances$state_name)]
```

#Cold war
```{r}
NATO <- c("United States of America", "United Kingdom", "Belgium", "Netherlands", "France","Canada", "Portugal", "Italy", "Norway", "Denmark", "Icealnd", "Turkey", "Greece")
nato_ccode  <- member_alliances$ccode[match(NATO, member_alliances$state_name)]

warsaw <- c("Russia", "Czechoslovakia", "Romania", "Bulgaria", "Poland", "Hungary")
warsaw_ccode  <- member_alliances$ccode[match(warsaw, member_alliances$state_name)]
```


#Korean War
```{r}
sk <- c("United States of America", "United Kingdom", "South Korea")
sk_ccode  <- member_alliances$ccode[match(sk, member_alliances$state_name)]


nk <- c("Russia", "North Korea", "China")
nk_ccode  <- member_alliances$ccode[match(nk, member_alliances$state_name)]

```

#Vietnam War

```{r}
us <- c("United States of America", "South Korea", "China", "New Zealand", "Australia")
us_ccode  <- member_alliances$ccode[match(us, member_alliances$state_name)]


viet <- c("Russia", "North Korea", "Vietnam", "Cuba")
viet_ccode  <- member_alliances$ccode[match(viet, member_alliances$state_name)]
```

#Afghanistan war

```{r}
us <- c("United States of America", "United Kingdom", "Canada", "Germany", "Australia")
us_ccode  <- member_alliances$ccode[match(us, member_alliances$state_name)]


afg <- c("Afghanistan")
afg_ccode  <- member_alliances$ccode[match(afg, member_alliances$state_name)]
```


```{r}

countries<- c("United States of America", "United Kingdom", "Russia", "Japan", "Italy","Germany", "Turkey", "Austria-Hungary", "Romania", "Bulgaria", "Canada", "Australia", "China", "Vietnam", "South Korea", "North Korea", "Poland", "Austria", "Hungary", "Greece")
ct_ccode  <- member_alliances$ccode[match(countries, member_alliances$state_name)]



```

```{r fig.width=6, fig.height=8}
reverse <- filter(dyad_al_year, ccode2 %in% ct_ccode)
reverse <- reverse[reverse$dyad_st_year >= 1900,]
forward <- filter(dyad_al, ccode1 %in% ct_ccode)

d=data.frame(x1=c(1914,1939, 1947, 1950, 1955, 2001), x2=c(1918, 1945, 1991, 1953, 1975, 2010), y1=c(0,0,0,0,0,0), y2=c(.4,.4,.4,.4,.4,.4), Conflict=c("WWI", "WWII", "Cold War", "Korean War", "Vietnam War", "Afghanistan War"), r=c(1,2,3,4,5,6))

ggplot() + 
  geom_point(data =reverse, aes(x=year, y = state_name2, color = state_name1)) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Australia", ymax="Vietnam", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom")


```

"United States of America", "United Kingdom", "Russia", "Japan", "Italy","Germany", "Turkey", "Austria-Hungary", "Romania", "Bulgaria", "Canada", "Australia", "China", "Vietnam", "South Korea", "North Korea", "Poland", "Austria", "Hungary", "Greece"

```{r fig.width=12, fig.height=12}

dir_alliances <- gather(dir_al_year, treaty_type, idicator, defense:entente)

dir_alliances <- dir_alliances[!dir_alliances$idicator %in%  0,]

dir_alliances$dyad_end_year[dir_alliances$dyad_end_year %in% NA] = 2016
dir_alliances <- dir_alliances[dir_alliances$year>1900,]

ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 == 2, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("United States Alliances")

```


```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 == 200, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("United Kingdom Alliances")

```

```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 == 255, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("Germany")

```


```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 == 220, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("France")

```

```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 == 365, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("Russia")

```


```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 == 700, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("Afghanistan")

```


```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 == 20, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("Canada")

```




```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 == 40, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("Cuba")

```





```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 == 750, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("India")

```






```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 ==325, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("Italy")

```


```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 ==740, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("Japan")

```



```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 ==731, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("North Korea")

```



```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 ==230, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("Spain")

```




```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 ==300, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("Austri-Hungary")

```



```{r fig.width=12, fig.height=12}
ggplot() + 
  geom_point(data =dir_alliances[dir_alliances$ccode1 ==816, ], aes(x=year, y = state_name2, color = treaty_type), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin="Afghanistan", ymax="Zimbabwe", fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("Vietnam")

```


```{r}

#number of alliances 

alliance_count <- dir_alliances[, c(2,3,14)]
alliance_count$count <- 1 

gp_ct <- aggregate(cbind(count) ~ ccode1+state_name1+year, data = alliance_count, FUN = sum )

#countries <- c("Afghanistan", "Austria-Hungary", "Canada", "Cuba", "France", "Germany", "India", "Italy", "Japan", "North Korea", "Russia", "Spain", "United Kingdom", "United States of America", "Vietnam")
countries <- c("China", "France", "Germany", "Japan",  "Russia", "United Kingdom", "United States of America")
countires_code <-  member_alliances$ccode[match(countries, member_alliances$state_name)]

gp_ct <- filter(gp_ct, ccode1 %in% countires_code)


```



```{r fig.width=14, fig.height=8}

ggplot() + 
  geom_line(data =gp_ct, aes(x=year, y = count, color = state_name1), alpha = .5) + 
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=0, ymax=max(gp_ct$count), fill=Conflict),alpha=0.15) + 
  theme(legend.position="bottom") + ggtitle("Number of Alliances by Year")
```

