---
title: "MID_final"
output:
  html_notebook: default
  pdf_document: default
---


## About the data - Militarized Interstate Disputes  
The data set records all instances of when one state threatened, displayed, or used force against another. 

```{r fig.width=12, fig.height=8}
library(ggplot2)
library(dplyr)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(GGally)

#path = "/home/vaguiar/col_hw/vis_hw/final/data/"
MIDA = read.csv(file="./data/MID/MIDA_4.01.csv", sep= ",")

war_year <- function(x){
  if(x >= 1914 & x <= 1918)
   return('WWI')
  if(x >= 1939 & x <= 1945)
   return('WWII')
  if(x >= 1950 & x <= 1953)
   return('Korean War')
  if(x >= 1955 & x <= 1975)
   return('Vietnam War')
  if(x >= 1947 & x <= 1991)
   return('Cold War')
  if(x >= 2001 & x <= 2010)
   return('War in Afghanistan')
  else
    return('No War')
}

MIDA$wartime <- sapply(MIDA$StYear, war_year)

##PCP PLot 
alphabending = 0.5
war <- ggparcoord(MIDA[MIDA$StYear>1900 & MIDA$wartime!='No War',], columns = c(9:12, 15:16), 
                  scale = "uniminmax", 
                  alphaLines = alphabending,
                  groupColumn = "wartime", 
                  title="Correlations In War Time Conflicts")
peace <- ggparcoord(MIDA[MIDA$StYear>1900 & MIDA$wartime=='No War',], columns = c(9:12, 14:16), 
                  scale = "uniminmax",
                  alphaLines = alphabending,
                  groupColumn = "wartime", 
                  title="Correlations In Peace Time Conflicts")
grid.arrange(war, peace, nrow = 2)  
```

The data is rich with information about the Outcomes, Settlements, The Number of Fatalities, Minimum Duration, The Highest Action taken and Hostility Level during each militarized conflict during that time period. We've selected a few important variables and plotted them in a pcp to see if we can spot a correlation between them.

Below, we dig a little deeper into a few variables like Fatalities, Settlements and Outcomes of the conflicts last century. Each graph is overlayed with the time period of major wars. Through this exploration, we try to spot possible correlations between the variables and the start of a war.

## Fatalaties In Disputes Leading to War

```{r fig.width=12, fig.height=10}

# Overlay w/ rectangle theme
y2_high =9
d = data.frame(x1=c(1914,1939, 1947, 1950, 1955, 2001), x2=c(1918, 1945, 1991, 1953, 1975, 2010), y1=c(0,0,0,0,0,0), y2=c(y2_high,y2_high,y2_high,y2_high,y2_high,y2_high), Conflict=c("WWI", "WWII", "Cold War", "Korean War", "Vietnam War", "Afghanistan War"), r=c(1,2,3,4,5,6))


#Using distinct color palette
cols <- colorRampPalette(brewer.pal(12, "Paired"))
myPal <- cols(length(unique(nat_top10_exp$statename)))

#Adding Labels for Facet titles
facet_names <- as_labeller(c(
  '0' = "None",
  '1' = "1-25 deaths",
  '2' = "26-100 deaths",
  '3' = "101-250 deaths",
  '4' = "251-500 deaths",
  '5' = "501-999 deaths",
  '6' = "More than 999 deaths",
  '-9' = "Missing Data"
))

ggplot() + 
  geom_bar(data=MIDA[MIDA$EndYear>1900 & MIDA$Fatality!=0 & MIDA$Fatality!=-9,], aes(x = EndYear),   stat="count") +
  facet_wrap(~Fatality, nrow= 8, labeller=facet_names) +
  scale_fill_manual(values = myPal) +
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2, fill=Conflict), alpha=0.15) +
  ggtitle("Number Of Disputes With Fatalities By Year") +
  xlab("End Year") +
  ylab("Count")

```
In the graph above we have plotted the Fatalites occuring due to militarized conflicts every year since the year 1900. We've faceted the graphs by the number of casualties starting from 1-25, 25-100 and so on.  
  
As we can see from the graphs above, there were conflicts resulting in more than 999 deaths in the years leading to the each and every major war. And in four of those years, namely 1913, 1938, 1955, and 2001, we see spikes in the number of deaths. This could be indicative of the fact that those conflicts forced countries to declare wars on each other.



** Explain Missing Data in the Fatalities

## Settlement of Disputes Leading to War

```{r fig.width=12, fig.height=6}

# Overlay w/ rectangle theme
y2_high = 9
d = data.frame(x1=c(1914,1939, 1947, 1950, 1955, 2001), x2=c(1918, 1945, 1991, 1953, 1975, 2010), y1=c(0,0,0,0,0,0), y2=c(y2_high,y2_high,y2_high,y2_high,y2_high,y2_high), Conflict=c("WWI", "WWII", "Cold War", "Korean War", "Vietnam War", "Afghanistan War"), r=c(1,2,3,4,5,6))


#Adding Labels for Facet titles
facet_names <- as_labeller(c(
  '1' = "Negotiated",
  '2' = "Imposed",
  '3' = "None",
  '4' = "Unclear",
  '-9' = "Missing Data"
))

ggplot() + 
  geom_bar(data=MIDA[MIDA$EndYear>1900 & MIDA$Settle!=3 & MIDA$Settle!='-9', ], aes(x = EndYear), stat="count") +
  facet_wrap(~Settle, nrow= 5, labeller=facet_names) +
  scale_fill_manual(values = myPal) +
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2, fill=Conflict), alpha=0.15) +
  ggtitle("Number Of Disputes With Settlement Outcomes By Year") +
  xlab("End Year") +
  ylab("Count")
```
In the graph above we have plotted the Settlement outcomes of different militarized conflicts every year since the year 1900. We've faceted the graphs by the Settlement types and for simplicity we've left out missing data or settlements that were unclear.  
  
Surprisingly, we don't see any correlations between Imposed/Negotiated Settlements and the war. But on second though, the number of Settlements, would not be an important predictor of war. It's the quality and the intricacies of Settlement, such as the Treaty of Versailles, that had the most impact on certain wars. The data doesn't delve into the qualitative aspects of these Settlements and hence it is of little or no significance to our exploration.

## Outcomes of Disputes Leading to War

```{r fig.width=12, fig.height=10}

# Overlay w/ rectangle theme
y2_high =65
d = data.frame(x1=c(1914,1939, 1947, 1950, 1955, 2001), x2=c(1918, 1945, 1991, 1953, 1975, 2010), y1=c(0,0,0,0,0,0), y2=c(y2_high,y2_high,y2_high,y2_high,y2_high,y2_high), Conflict=c("WWI", "WWII", "Cold War", "Korean War", "Vietnam War", "Afghanistan War"), r=c(1,2,3,4,5,6))

MIDA_o <- filter(MIDA, MIDA$Outcome %in% c(1,2,3,4,5,6,8,-9))
war_outcome <- function(x){
  if(x ==2)
   return(1)
  if(x == 4)
   return(3)
  else
    return(x)
}

MIDA_o$WarOutcome <- sapply(MIDA_o$Outcome, war_outcome)


#Adding Labels for Facet titles
facet_names <- as_labeller(c(
  '1' = "Victory For Either Side",
  '3' = "Yield By Either Side",
  '5' = "Stalemate",
  '6' = "Compromise", 
  '8' = "Unclear",
  '-9' = "Missing Data"
))

ggplot() + 
  geom_bar(data=MIDA_o[MIDA_o$EndYear>1900 & MIDA_o$WarOutcome!='-9' , ], 
  aes(x = EndYear), stat="count") +
  facet_wrap(~WarOutcome, nrow= 5, labeller=facet_names) +
  scale_fill_manual(values = myPal) +
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2, fill=Conflict), alpha=0.15) +
  ggtitle("Number Of Disputes With Outcomes By Year") +
  xlab("End Year") +
  ylab("Count")
```

In the graph above we have plotted the Outcomes of different militarized conflicts every year since the year 1900. We've faceted the graphs by the Outcomes and for simplicity, we've left out missing data or outcomes that were unclear.  

There is little evidence to show that certain types Outcomes were predictive of a war. But surprisingly so, we see that Stalemates peaked during the middle of an ongoing war. We see that in WWII, the Vietnam War and during War in Afghanistan.  
  
  

#### Viewing Initiators/Victims of most conflicts 

```{r fig.width=12, fig.height=12}
library(ggrepel)

MIDB = read.csv(file="./data/MID/MIDB_4.01.csv", sep= ",")

#Filter by countries that originated disputes.
MIDB_O <- filter(MIDB, Orig==1)

#Filter by countries that attacked during various disputes.
MIDB_A <- filter(MIDB, Orig==0)

#summary(MIDB_A)

#Count the conflicts originated by countries each year
WarOrig <- MIDB_O %>% count(StAbb, sort=TRUE, vars=StYear)

#Count the conflicts in which countries were attacked each year
WarAttack <- MIDB_A %>% count(StAbb, sort=TRUE, vars=StYear)

#Top Conflicts stared  Countries by Year
for(yr in 1901:2010)
{
  WarOrig_yr <- head(arrange(filter(WarOrig, vars==yr), desc(n)), 1)
  WarAttack_yr <- head(arrange(filter(WarAttack, vars==yr), desc(n)), 1)
    
  if(yr == 1901)
  {
    WarTopOrig <- WarOrig_yr
    WarTopAttacked <- WarAttack_yr
  }
  else
  {
    WarTopOrig <- rbind(WarTopOrig, WarOrig_yr)
    WarTopAttacked <- rbind(WarTopAttacked, WarAttack_yr)
  }
}

#print(WarTopOrig)

# Overlay w/ rectangle theme
y2_high =30
d = data.frame(x1=c(1914,1939, 1947, 1950, 1955, 2001), x2=c(1918, 1945, 1991, 1953, 1975, 2010), y1=c(0,0,0,0,0,0), y2=c(y2_high,y2_high,y2_high,y2_high,y2_high,y2_high), Conflict=c("WWI", "WWII", "Cold War", "Korean War", "Vietnam War", "Afghanistan War"), r=c(1,2,3,4,5,6))

confictsInitiated = 6
orig <- ggplot() + 
  geom_point(data=WarTopOrig, aes(x = vars, y = n), size = 1, color = 'red') +
  geom_label_repel(data=WarTopOrig,
    aes(x = vars, y = n, label= ifelse(n>=confictsInitiated, as.character(StAbb), ''), 
        fill = factor(StAbb)),
    fontface = 'bold', color = 'white',
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.color = 'grey50') +
    theme(legend.position = "none") +
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2, fill=Conflict), alpha=0.15) +
  theme_classic(base_size = 16) +
  ggtitle("Top Countries Starting Conflicts") +
  xlab("Year") +
  ylab("Count") + theme(legend.position = "none")

timesAttacked = 2
yNew = 11
attacked <- ggplot() + 
  geom_point(data=WarTopAttacked, aes(x = vars, y = n), size = 1, color = 'red') +
  geom_label_repel(data=WarTopAttacked,
    aes(x = vars, y = n, label= ifelse(n>=timesAttacked, as.character(StAbb), ''), 
        fill = factor(StAbb)),
    fontface = 'bold', color = 'white',
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.color = 'grey50') +
    theme(legend.position = "none") +
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=yNew, fill=Conflict), alpha=0.15) +
  theme_classic(base_size = 16) +
  ggtitle("Countries Most Attacked") +
  xlab("Year") +
  ylab("Count") + theme(legend.position = "none")

grid.arrange(orig, attacked, nrow = 2)
```

The purpose of the graph above was to plot the most belligerent/attacked countries during conflicts happening each year.

The graph sheds light on the number of conflicts happening before, during and towards the end of a war. In quite a few cases the top initiators of conflicts had heavy involvement in the war.
