---
title: "MID_final"
output:
  html_notebook: default
  pdf_document: default
---


## About the data - Militarized Interstate Disputes  
The data set records all instances of when one state threatened, displayed, or used force against another. 

```{r fig.width=12, fig.height=8}
library(ggplot2)
library(dplyr)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(GGally)

path = "/home/vaguiar/col_hw/vis_hw/final/data/"
MIDA = read.csv(file="./data/MID/MIDA_4.01.csv", sep= ",")
MIDB = read.csv(file="./data/MID/MIDB_4.01.csv", sep= ",")

war_year <- function(x){
  if(x >= 1914 & x <= 1918)
   return('WWI')
  if(x >= 1939 & x <= 1945)
   return('WWII')
  if(x >= 1950 & x <= 1953)
   return('Korean War')
  if(x >= 1955 & x <= 1975)
   return('Vietnam War')
  if(x >= 1947 & x <= 1991)
   return('Cold War')
  if(x >= 2001 & x <= 2010)
   return('War in Afghanistan')
  else
    return('No War')
}

MIDA$wartime <- sapply(MIDA$StYear, war_year)

##PCP PLot 
alphabending = 0.5
war <- ggparcoord(MIDA[MIDA$StYear>1900 & MIDA$wartime!='No War',], columns = c(9:12, 15:16), 
                  scale = "uniminmax", 
                  alphaLines = alphabending,
                  groupColumn = "wartime", 
                  title="Correlations In War Time Conflicts")
peace <- ggparcoord(MIDA[MIDA$StYear>1900 & MIDA$wartime=='No War',], columns = c(9:12, 14:16), 
                  scale = "uniminmax",
                  alphaLines = alphabending,
                  groupColumn = "wartime", 
                  title="Correlations In Peace Time Conflicts")
grid.arrange(war, peace, nrow = 2)  
```

The data is rich with information about the Outcomes, Settlements, The Number of Fatalities, Minimum Duration, The Highest Action taken and Hostility Level during each militarized conflict during theat time period. We've selected a few important variables and plotted them in a pcp to see the correlation between them.

Below we dig a little deeper into a few variables like Fatalities, Settlements and Outcomes of the conflicts last century. Each graph is overlayed with the time period of major wars. Through this exploration, we try to spot possible correlations between the variables and the start of a war.

## Fatalaties In Disputes Leading to War

```{r fig.width=12, fig.height=10}

# Overlay w/ rectangle theme
y2_high =9
d = data.frame(x1=c(1914,1939, 1947, 1950, 1955, 2001), x2=c(1918, 1945, 1991, 1953, 1975, 2010), y1=c(0,0,0,0,0,0), y2=c(y2_high,y2_high,y2_high,y2_high,y2_high,y2_high), Conflict=c("WWI", "WWII", "Cold War", "Korean War", "Vietnam War", "Afghanistan War"), r=c(1,2,3,4,5,6))


#Using distinct color palette
cols <- colorRampPalette(brewer.pal(12, "Paired"))
myPal <- cols(length(unique(nat_top10_exp$statename)))

#Adding Labels for Facet titles
facet_names <- as_labeller(c(
  '0' = "None",
  '1' = "1-25 deaths",
  '2' = "26-100 deaths",
  '3' = "101-250 deaths",
  '4' = "251-500 deaths",
  '5' = "501-999 deaths",
  '6' = "More than 999 deaths",
  '-9' = "Missing Data"
))

ggplot() + 
  geom_bar(data=MIDA[MIDA$EndYear>1900 & MIDA$Fatality!=0 & MIDA$Fatality!=-9,], aes(x = EndYear),   stat="count") +
  facet_wrap(~Fatality, nrow= 8, labeller=facet_names) +
  scale_fill_manual(values = myPal) +
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2, fill=Conflict), alpha=0.15) +
  ggtitle("Number Of Disputes With Fatalities By Year") +
  xlab("End Year") +
  ylab("Count")

```


** Explain Missing Data in the Fatalities

## Settlement of Disputes Leading to War

```{r fig.width=12, fig.height=6}

# Overlay w/ rectangle theme
y2_high = 9
d = data.frame(x1=c(1914,1939, 1947, 1950, 1955, 2001), x2=c(1918, 1945, 1991, 1953, 1975, 2010), y1=c(0,0,0,0,0,0), y2=c(y2_high,y2_high,y2_high,y2_high,y2_high,y2_high), Conflict=c("WWI", "WWII", "Cold War", "Korean War", "Vietnam War", "Afghanistan War"), r=c(1,2,3,4,5,6))


#Adding Labels for Facet titles
facet_names <- as_labeller(c(
  '1' = "Negotiated",
  '2' = "Imposed",
  '3' = "None",
  '4' = "Unclear",
  '-9' = "Missing Data"
))

ggplot() + 
  geom_bar(data=MIDA[MIDA$EndYear>1900 & MIDA$Settle!=3 & MIDA$Settle!='-9', ], aes(x = EndYear), stat="count") +
  facet_wrap(~Settle, nrow= 5, labeller=facet_names) +
  scale_fill_manual(values = myPal) +
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2, fill=Conflict), alpha=0.15) +
  ggtitle("Number Of Disputes With Settlement Outcomes By Year") +
  xlab("End Year") +
  ylab("Count")
```

## Outcomes of Disputes Leading to War

```{r fig.width=12, fig.height=10}

# Overlay w/ rectangle theme
y2_high =65
d = data.frame(x1=c(1914,1939, 1947, 1950, 1955, 2001), x2=c(1918, 1945, 1991, 1953, 1975, 2010), y1=c(0,0,0,0,0,0), y2=c(y2_high,y2_high,y2_high,y2_high,y2_high,y2_high), Conflict=c("WWI", "WWII", "Cold War", "Korean War", "Vietnam War", "Afghanistan War"), r=c(1,2,3,4,5,6))

MIDA_o <- filter(MIDA, MIDA$Outcome %in% c(1,2,3,4,5,6,8,-9))
war_outcome <- function(x){
  if(x ==2)
   return(1)
  if(x == 4)
   return(3)
  else
    return(x)
}

MIDA_o$WarOutcome <- sapply(MIDA_o$Outcome, war_outcome)


#Adding Labels for Facet titles
facet_names <- as_labeller(c(
  '1' = "Victory For Either Side",
  '3' = "Yield By Either Side",
  '5' = "Stalemate",
  '6' = "Compromise", 
  '8' = "Unclear",
  '-9' = "Missing Data"
))

ggplot() + 
  geom_bar(data=MIDA_o[MIDA_o$EndYear>1900 & MIDA_o$WarOutcome!='-9' , ], 
  aes(x = EndYear), stat="count") +
  facet_wrap(~WarOutcome, nrow= 5, labeller=facet_names) +
  scale_fill_manual(values = myPal) +
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2, fill=Conflict), alpha=0.15) +
  ggtitle("Number Of Disputes With Outcomes By Year") +
  xlab("End Year") +
  ylab("Count")
```
